Quasars Probing Quasars IX. The Kinematics of the Circumgalactic Medium Surrounding $z\sim2$ Quasars [v1.4]

### 2017-03-24 QPQ9/qpq9_draft.tex
Put the noteobok for making the QPQ9 structure in a module Small changes

\begin{abstract}
We examine the kinematics of the gas in the environments of galaxies hosting quasars at $z\sim2$. 
We employ 112 projected quasar pairs to study the circumgalactic gas of the foreground quasars in 
absorption. The sample selects foreground quasars with precise redshift measurements, using 
emission-lines with precision $\lesssim300\,{\rm km\,s^{-1}}$ and average offsets from the 
systemic redshift $\lesssim|100\,{\rm km\,s^{-1}}|$. We stack the background quasar spectra at the 
foreground quasar's systemic redshift to study the mean absorption in \ion{C}{2}, \ion{C}{4}, 
and \ion{Mg}{2}. We find that the mean absorptions exhibit large velocity widths 
$\sigma_v>300\,{\rm km\,s^{-1}}$. We show that, gas in gravitational motion, Hubble flow, or 
galactic-scale outflows are all consistent with the observed widths. Furthermore, we find that the 
mean absorptions are asymmetric about the systemic redshift. The mean absorption centroids exhibit 
small redshift relative to the systemic $\delta v\approx+200\,{\rm km\,s^{-1}}$, with large 
intrinsic scatter in the centroid velocities of the individual absorption systems. We show 
that the observed offsets may be produced by anisotropic or intermittent ionizing radiation from 
the foreground quasars. 
\end{abstract}

@@

\begin{figure}
\includegraphics[width=3.5in]{Figures/fig_stacks_and_fits.pdf}
%\includegraphics[width=3.5in]{f02.pdf}
\caption{Mean and median absorption centered at \ion{C}{2}\,1334, \ion{C}{4}\,1548, and 
\ion{Mg}{2}\,2796 of the foreground quasars for all QPQ9 pairs. The composites are shown in thick, 
black. Overplotted on the composites are Gaussian fits, normalized to pseudo-continua far away 
from a velocity of $0\,{\rm km\,s^{-1}}$ relative to $z_{\rm fg}$, and are shown in thin, blue. 
For the doublets, a second Gaussian with a fixed mean separation and a tied standard deviation is 
included in the modeling. The absorptions frequently exhibit large velocity widths. The dashed 
lines mark the centroids, which show small positive velocity offsets from $z_{\rm fg}$. The 
$1\sigma$ modeling error of the centroid and dispersion of the \ion{C}{2} mean stack is 
$\approx35\,{\rm km\,s^{-1}}$, which is also the typical modeling errors of the other stacks. 
}
\label{fig:stacks_and_fits}
\end{figure}

@@

\begin{figure}
\includegraphics[width=2.5in]{Figures/fig_monte.pdf}
%\includegraphics[width=2.5in]{f05.pdf}
\caption{The same \ion{C}{2} mean stack shown in the first panel of 
Figure~\ref{fig:stacks_and_fits} is shown in thick, black. In thin, green, we overplot the 
Gaussian absorption model of the Monte Carlo simulations generated from a purely clustering 
argument. The model from clustering analysis is multiplied to the pseudo-continuum level of the 
stack of the observational data, shifted to the centroid of the stack of the data, and broadened 
by the mean redshift error in the data. The model is a good description of the data, with a 
dispersion within modeling error of the dispersion in the data. We overplot in dashed, yellow an 
absorption profile with $\sigma_v=654\,{\rm km\,s^{-1}}$, that is ruled out. Motions that produce 
this width are ruled out at $3\sigma$ confidence by bootstrap analysis on the observational data. 
}
\label{fig:monte}
\end{figure}

@@

\cite{Eftekharzadeh+15} measured the clustering of quasars in the range $2.2<z<2.8$ while 
\cite{RodriguezTorres+16} measured the clustering of quasars in the range $1.8<z<2.2$. They 
estimated that these quasars are hosted by dark matter halos with mass $10^{12.5}\,M_\odot$ and 
$10^{12.6}\,M_\odot$ respectively. If QPQ dark matter halos have a characteristic mass 
$10^{12.5}\,M_\odot$ and follow an NFW profile \citep{NavarroFrenkWhite97} with concentration 
parameter $c=4$, at $z\approx2$ the maximum circular velocity is $319\,{\rm km\,s^{-1}}$. 
According to \cite{Tormen97}, the maximum circular velocity is $\approx1.4$ times larger than the 
one-dimensional velocity dispersion $\sigma_{\rm 1D}$. Hence, the line-of-sight velocity 
dispersion typical of QPQ halos is $\sigma_{\rm 1D}=228\,{\rm km\,s^{-1}}$. We can investigate 
whether gravitational motions and Hubble flows are sufficient to reproduce the dispersion in the 
data, using Monte Carlo methods to simulate the absorption signals.

Since \ion{C}{2} systems arise in optically thick absorbers, we may adopt the clustering analysis 
results of QPQ6. In the absence of clustering, the expected number of absorbers per unit redshift 
interval for Lyman limit systems, super Lyman limit systems, and damped Ly$\alpha$ systems are 
respectively $\ell_{\rm IGM}^{\rm LLS}(z)\approx1.05((1+z)/(1+2.5))^{2.1}$, 
$\ell_{\rm IGM}^{\rm SLLS}(z)\approx0.44((1+z)/(1+2.5))^{2.1}$, and 
$\ell_{\rm IGM}^{\rm DLA}(z)\approx0.2((1+z)/(1+2.5))^{2.1}$. The quasar-absorber correlation 
functions for Lyman limit systems, super Lyman limit systems, and damped Ly$\alpha$ systems are 
respectively $\xi_{\rm QA}^{\rm LLS}(r)=(r/(12.5\,h^{-1}\,{\rm Mpc}))^{-1.68}$, 
$\xi_{\rm QA}^{\rm SLLS}(r)=(r/(14.0\,h^{-1}\,{\rm Mpc}))^{-1.68}$, and 
$\xi_{\rm QA}^{\rm DLA}(r)=(r/(3.9\,h^{-1}\,{\rm Mpc}))^{-1.6}$. 
For each quasar pair, we calculate the expected number of optically thick absorbers within 
$\pm3000\,{\rm km\,s^{-1}}$ at a distance $R_\perp$ from the foreground quasar and at 
$z_{\rm fg}$. Then we generate 1000 mock sightlines. The number of absorbers for each mock 
spectrum is randomly selected from a Poisson distribution with mean equal to the expected number 
calculated as above. The absorbers are randomly assigned Hubble velocities, with a probability 
distribution according to the quasar-absorber correlation functions. The absorbers are randomly 
assigned additional peculiar velocities drawn from a normal distribution with mean equal to 
$0\,{\rm km\,s^{-1}}$ and scatter equal to $\sigma_{\rm 1D}$. For each absorber, we assume a rest 
equivalent width for \ion{C}{2} $W_{\rm CII}$ and a Gaussian absorption profile. We repeat the 
above procedure for all 32 quasar pairs, and create a mean stack of the 32000 mock spectra 
generated. We fit a Gaussian absorption profile multiplied to a constant continuum level to model 
the stack of mock spectra. We adjust the $W_{\rm CII}$ adopted for the absorbers until the 
amplitude of the best-fit Gaussian of the stack of mock spectra matches the 
amplitude of the stack of the observational data. We find that $W_{\rm CII}=0.7\,{\rm \AA}$ well 
reproduces the amplitude, and the dispersion of the Gaussian absorption model is 
insensitive to the assumed $W_{\rm CII}$ or line profile for one absorber.

In Figure~\ref{fig:monte}, we show a comparison of the observational data stack and the Gaussian 
absorption model of the Monte Carlo simulations. In the figure, the Gaussian absorption model is 
broadened by the mean redshift error by adding it in quadrature to the dispersion in the model. 
The resulting stack of mock spectra has a $1\sigma$ dispersion of $263\,{\rm km\,s^{-1}}$. This is 
within $1\sigma$ modeling error of the intrinsic dispersion in the \ion{C}{2} mean stack for the 
full sample, and about 2 times the modeling error away from the intrinsic dispersion in the stack 
of the sub-sample with [\ion{O}{3}] redshifts. We also test for the 
sensitivity of this measured dispersion to the correlation functions adopted. The QPQ6 clustering 
analysis is performed on only the strongest absorber near $z_{\rm fg}$, and a low $R_\perp$ 
sightline may in fact intercept more than one optically thick absorber. We double the number of 
absorbers for each mock sightline, and find the measured dispersion only increases by several 
${\rm km\,s^{-1}}$. Thus, the hypothesis that the observed velocity width is only produced by a 
combination of gravitational motions and Hubble flows cannot be ruled out. Extra dynamical 
processes are not necessary to explain the large velocity fields. From the bootstrap analysis of 
the observed dispersion, a width of $\sigma_v>654\,{\rm km\,s^{-1}}$ is ruled out at 3$\sigma$ 
confidence (overplotted in Figure~\ref{fig:monte}). In other words, motions additional to 
gravitational and Hubble flows that will produce intrinsic dispersion 
$\sigma_v>628\,{\rm km\,s^{-1}}$ are ruled out.

On the contrary, $\sigma_{\rm 1D}<175\,{\rm km\,s^{-1}}$ together with Hubble velocities will 
result in a velocity width that is more than three times the modeling error away from the observed 
intrinsic width. This implies that if the characteristic dark matter halo mass 
$<10^{12.2}\,M_\odot$, additional dynamical processes are required to explain the observed width.

@@

As a test of reasonableness, we calculate that, at $R_\perp\approx50\textendash150$\,kpc, radial 
speeds relative to the foreground quasar of $500\textendash700\,{\rm km\,s^{-1}}$ and quasar 
opening angles of $120^\circ\textendash140^\circ$ may reproduce the observed centroid and 
intrinsic dispersion of the \ion{C}{2} mean absorption. The first explanation above to asymmetric 
absorption requires that the quasars emit their ionizing radiation anisotropically, while the 
second explanation requires the quasars emit their ionizing radiation intermittently. Both 
explanations will require the gas to be either outflowing or in Hubble flow. Were the gas 
inflowing instead, the velocity centroid would be negative.

>>>>

\begin{abstract}
We examine the kinematics of the gas in the environments of galaxies hosting quasars at $z\sim2$. 
We employ 112 projected quasar pairs to study the circumgalactic gas of the foreground quasars in 
absorption. The sample selects foreground quasars with precise redshift measurements, using 
emission-lines with precision $\lesssim300\,{\rm km\,s^{-1}}$ and average offsets from the 
systemic redshift $\lesssim|100\,{\rm km\,s^{-1}}|$. We stack the background quasar spectra at the 
foreground quasar's systemic redshift to study the mean absorption in \ion{C}{2}, \ion{C}{4}, 
and \ion{Mg}{2}. We find that the mean absorptions exhibit large velocity widths 
$\sigma_v>300\,{\rm km\,s^{-1}}$. The observed widths are consistent with gas in gravitational 
motion and Hubble flow, and galactic-scale outflows are not be required to explain the large 
widths. Furthermore, we find that the mean absorptions are asymmetric about the systemic redshift. 
The mean absorption centroids exhibit small redshift relative to the 
systemic $\delta v\approx+200\,{\rm km\,s^{-1}}$, with large intrinsic scatter in the centroid 
velocities of the individual absorption systems. We show that the observed offsets may be produced 
if (i) the ionizing radiation from the foreground quasars is anisotropic or intermittent; (ii) the 
gas is not flowing into the galaxy.  
\end{abstract}

@@

\begin{figure}
\includegraphics[width=3.5in]{Figures/fig_stacks_and_fits.pdf}
%\includegraphics[width=3.5in]{f02.pdf}
\caption{Mean and median absorption centered at \ion{C}{2}\,1334, \ion{C}{4}\,1548, and 
\ion{Mg}{2}\,2796 of the foreground quasars for all QPQ9 pairs. The composites are shown in thin, 
black. Overplotted on the composites are Gaussian fits, normalized to pseudo-continua far away 
from a velocity of $0\,{\rm km\,s^{-1}}$ relative to $z_{\rm fg}$, and are shown in thick, blue. 
For the doublets, a second Gaussian with a fixed mean separation and a tied standard deviation is 
included in the modeling. The absorptions frequently exhibit large velocity widths. The dashed 
lines mark the centroids, which show small positive velocity offsets from $z_{\rm fg}$. The 
$1\sigma$ modeling error of the centroid and dispersion of the \ion{C}{2} mean stack is 
$\approx35\,{\rm km\,s^{-1}}$, which is also the typical modeling errors of the other stacks. 
}
\label{fig:stacks_and_fits}
\end{figure}

@@

\begin{figure}
\includegraphics[width=2.5in]{Figures/fig_monte.pdf}
%\includegraphics[width=2.5in]{f05.pdf}
\caption{The same \ion{C}{2} mean stack shown in the first panel of 
Figure~\ref{fig:stacks_and_fits} is shown in thin, black. In thick, green, we overplot the 
Gaussian absorption model of the Monte Carlo simulations generated from a purely clustering 
argument. The model from clustering analysis is multiplied to the pseudo-continuum level of the 
stack of the observational data, shifted to the centroid of the stack of the data, and broadened 
by the mean redshift error in the data. The model is a good description of the data, with a 
dispersion within modeling error of the dispersion in the data. We overplot in dashed, yellow an 
absorption profile with $\sigma_v=654\,{\rm km\,s^{-1}}$, which is larger than the observed width 
by three times the standard deviation in the bootstrap analysis. Motions that produce a velocity 
width larger than this can be ruled out. We overplot in dashed, magenta an absorption profile with 
$\sigma_v=207\,{\rm km\,s^{-1}}$, which is smaller than the observed width by three times the 
modeling error. Unless gravitational and Hubble flows together with redshift error broadening 
produce a velocity width smaller than this, extra dynamical processes (e.g.\ outflows) will not be 
required to explain the observation. 
}
\label{fig:monte}
\end{figure}

@@

\cite{Eftekharzadeh+15} measured the clustering of quasars in the range $2.2<z<2.8$ while 
\cite{RodriguezTorres+16} measured the clustering of quasars in the range $1.8<z<2.2$. They 
estimated that these quasars are hosted by dark matter halos with mass $10^{12.5}\,M_\odot$ and 
$10^{12.6}\,M_\odot$ respectively. If dark matter halos hosting QPQ9 quasars have a characteristic 
mass $10^{12.6}\,M_\odot$ and follow an NFW profile \citep{NavarroFrenkWhite97} with concentration 
parameter $c=4$, at $z\approx2$ the maximum circular velocity is $345\,{\rm km\,s^{-1}}$. 
\cite{TormenBouchetWhite97} found that the maximum circular velocity is $\approx1.4$ times the 
maximum of the one-dimensional velocity dispersion $\sigma_{\rm 1D}$. Hence, the line-of-sight 
velocity dispersion typical of QPQ9 halos is $\sigma_{\rm 1D}=246\,{\rm km\,s^{-1}}$. We can 
investigate whether gravitational motions and Hubble flows are sufficient to reproduce the 
dispersion in the data, using Monte Carlo methods to simulate the absorption signals.

Since \ion{C}{2} systems arise in optically thick absorbers, we may adopt the clustering analysis 
results of QPQ6. In the absence of clustering, the expected number of absorbers per unit redshift 
interval for Lyman limit systems, super Lyman limit systems, and damped Ly$\alpha$ systems are 
respectively $\ell_{\rm IGM}^{\rm LLS}(z)\approx1.05((1+z)/(1+2.5))^{2.1}$, 
$\ell_{\rm IGM}^{\rm SLLS}(z)\approx0.44((1+z)/(1+2.5))^{2.1}$, and 
$\ell_{\rm IGM}^{\rm DLA}(z)\approx0.2((1+z)/(1+2.5))^{2.1}$. The quasar-absorber correlation 
functions for Lyman limit systems, super Lyman limit systems, and damped Ly$\alpha$ systems are 
respectively $\xi_{\rm QA}^{\rm LLS}(r)=(r/(12.5\,h^{-1}\,{\rm Mpc}))^{-1.68}$, 
$\xi_{\rm QA}^{\rm SLLS}(r)=(r/(14.0\,h^{-1}\,{\rm Mpc}))^{-1.68}$, and 
$\xi_{\rm QA}^{\rm DLA}(r)=(r/(3.9\,h^{-1}\,{\rm Mpc}))^{-1.6}$. 
For each quasar pair, we calculate the expected number of optically thick absorbers within 
$\pm3000\,{\rm km\,s^{-1}}$ at a distance $R_\perp$ from the foreground quasar and at 
$z_{\rm fg}$. Then we generate 1000 mock sightlines. The number of absorbers for each mock 
spectrum is randomly selected from a Poisson distribution with mean equal to the expected number 
calculated as above. The absorbers are randomly assigned Hubble velocities, with a probability 
distribution according to the quasar-absorber correlation functions. The absorbers are randomly 
assigned additional peculiar velocities drawn from a normal distribution with mean equal to 
$0\,{\rm km\,s^{-1}}$ and scatter equal to $\sigma_{\rm 1D}$. For each absorber, we assume a rest 
equivalent width for \ion{C}{2} $W_{\rm CII}$ and a Gaussian absorption profile. We repeat the 
above procedure for all 32 quasar pairs, and create a mean stack of the 32000 mock spectra 
generated. We fit a Gaussian absorption profile multiplied to a constant continuum level to model 
the stack of mock spectra. We adjust the $W_{\rm CII}$ adopted for the absorbers until the 
amplitude of the best-fit Gaussian of the stack of mock spectra matches the 
amplitude of the stack of the observational data. We find that $W_{\rm CII}=0.8\,{\rm \AA}$ well 
reproduces the amplitude, and the dispersion of the Gaussian absorption model is 
insensitive to the assumed $W_{\rm CII}$ or line profile for one absorber.

In Figure~\ref{fig:monte}, we show a comparison of the observational data stack and the Gaussian 
absorption model of the Monte Carlo simulations. In the figure, the Gaussian absorption model is 
broadened by the mean redshift error by adding it in quadrature to the dispersion in the model. 
The resulting stack of mock spectra has a $1\sigma$ dispersion of $281\,{\rm km\,s^{-1}}$. This is 
within $1\sigma$ modeling error from the intrinsic dispersion in the \ion{C}{2} mean stack for the 
full sample, and about 2 times the modeling error away from the intrinsic dispersion in the stack 
of the sub-sample with [\ion{O}{3}] redshifts. We also test for the 
sensitivity of this measured dispersion to the correlation functions adopted. The QPQ6 clustering 
analysis is performed on only the strongest absorber near $z_{\rm fg}$, and a low $R_\perp$ 
sightline may in fact intercept more than one optically thick absorber. We double the number of 
absorbers for each mock sightline, and find the measured dispersion only increases by several 
${\rm km\,s^{-1}}$. Thus, the hypothesis that the observed velocity width is only produced by a 
combination of gravitational motions and Hubble flows cannot be ruled out. Extra dynamical 
processes are not necessary to explain the large velocity fields. From the bootstrap analysis of 
the observed dispersion, a width of $\sigma_v>654\,{\rm km\,s^{-1}}$ would be larger than the 
observed by three times the scatter (overplotted in Figure~\ref{fig:monte}). In other words, 
taking into account the redshift errors, motions in addition to gravitational and Hubble 
flows that will produce an intrinsic dispersion $\sigma_v>628\,{\rm km\,s^{-1}}$ can be ruled out.

On the contrary, $\sigma_{\rm 1D}<74\,{\rm km\,s^{-1}}$ together with Hubble velocities and 
broadening by redshift errors will result in a velocity width that is more than three times the 
modeling error away from the observed width (overplotted in Figure~\ref{fig:monte}). This implies 
that, unless the characteristic dark matter halo mass $<10^{11.0}\,M_\odot$, additional dynamical 
processes are not required to explain the observed width.

@@

As a test of reasonableness, we calculate that, at $R_\perp\approx50\textendash150$\,kpc, radial 
speeds relative to the foreground quasar of $500\textendash700\,{\rm km\,s^{-1}}$ and quasar 
opening angles of $120^\circ\textendash140^\circ$ may reproduce the observed centroid and 
intrinsic dispersion of the \ion{C}{2} mean absorption. The first explanation above to asymmetric 
absorption requires that the quasars emit their ionizing radiation anisotropically, while the 
second explanation requires the quasars emit their ionizing radiation intermittently. Both 
explanations will require the line-of-sight motions of the gas to be in either outflow or Hubble 
flow. Were the gas flowing into the galaxy instead, the velocity centroid would be negative.
